---
title: "wrangle-data"
format: html
editor: visual
---

### Load in data

Here, I load in all the required packages to memory. Then, I read in all the files in a directory and store them in a list.

```{r load data}
#| include: false

library(tidyverse)
library(readxl)
library(sf)
library(ClimateNAr)
library(elevatr)
library(withr)
library(ncdf4)

# define the function to load all .xlsx files in a folder and put them in a list
load_xlsx_files <- function(directory) {
  # get the list of all .xlsx files in the specified directory
  file_list <- list.files(path = directory,
                          pattern = "*.xlsx",
                          full.names = TRUE)
  # extract file names without directory and extension, and clean them
  file_names <- str_replace_all(basename(file_list), 
                                "\\.[xX][lL][sS][xX]$", "")
  cleaned_names <- str_replace_all(file_names, "[^0-9]", "")
  
  # read each file and store the data in a list using purrr::map
  data_list <- purrr::map(file_list, read_excel)
  
  # assign cleaned names to the list elements
  names(data_list) <- cleaned_names
  
  # return the list of data frames
  return(data_list)
}

# run function to load in data
davis <- load_xlsx_files("../01_raw-data/davis/")
stebs <- load_xlsx_files("../01_raw-data/stebs/")

```

### Wrangle data frames

I wrangled the data matrix into observations by date (long format). I also added code to handle some of the idiosyncrasies in the spreadsheet format that caused errors when I tried to read it in.

```{r wrangle data}

# function to fill NA values with the most recent non-NA value in a row
fill_row <- function(row) {
  non_na_value <- NA
  for (i in 1:length(row)) {
    if (!is.na(row[i])) {
      non_na_value <- row[i]
    } else {
      row[i] <- non_na_value
    }
  }
  return(row)
}

# define the wrangle_data function for davis
wrangle_data_davis <- function(df, year) {
  # convert to df
  df <- df %>% as.data.frame()
  
  # remove the first row if there is a running total
  if (!is.na(df[3, 8]) && df[3, 8] == "Adults") {
    new_names <- ifelse(is.na(df[1, ]), paste0("...", seq_along(df[1, ])), df[1, ])
    new_names <- make.names(new_names, unique = TRUE)
    df <- setNames(df, new_names)
    df <- df %>% slice(-1)
  }
  
  # if night/month rows are flip-flopped
  if (df[1, 10] %in% c("N", "D")) {
    df[1, ] <- colnames(df) # colnames to first row
    df <- df %>% # remove day time data
      select(-which(str_detect(df[1, ], "D")))
    df[1, ] <- df[1, ] %>% # ... to NA
      str_replace(., "\\.\\.\\.", NA_character_) %>% as.list()
  }
  
  # apply the function only to the first row
  df[1, ] <- fill_row(as.vector(df[1, ]))
  
  # make a row for the dates
  date_row <- df %>%
    slice(c(1, 2)) %>%
    mutate_all(~ paste(., collapse = " "))
  
  # add it in
  df[1, ] <- date_row[1, ]
  
  # rename the columns as the first row
  df <- df %>%
    setNames(unlist(df[1, ]))
  
  # remove the last 4 columns
  df <- df[, -c((ncol(df) - 3):ncol(df))]
  
  # remove columns 3 through 6, 8, and 9
  df <- df %>%
    select(-3:-6, -8, -9)
  
  # define new column names for the species columns
  new_names <- c("hodges", "list", "verbatim_name")
  
  # small edits add up :)
  df <- df %>%
    # rename the species columns
    rename_with(~ new_names, c(1, 2, 3)) %>%
    # remove the first 2 rows using dplyr
    slice(-(1:2)) %>%
    # drop columns where "verbatim_name" is NA
    drop_na(verbatim_name) %>%
    # remove extra text
    filter(
      !verbatim_name == "Species:",
      !verbatim_name == "species (total)",
      !verbatim_name == "Individuals:",
      !verbatim_name == "Off:",
      !verbatim_name == "On:",
      !verbatim_name == "Temp., deg. F.:",
      !verbatim_name == "Hi Temp",
      !verbatim_name == "Lo Temp"
    ) %>%
    # remove periods from names
    rename_with(~ gsub("\\.", "", .), contains(".")) %>%
    # combine the first two columns into a single column
    mutate(hodges_list = ifelse(is.na(list), hodges, paste0(hodges, ".", list))) %>%
    # remove columns that were just used
    select(-c(hodges, list)) %>%
    # put hodges_list first
    select(hodges_list, everything()) %>%
    # pivot dates longer
    pivot_longer(
      cols = -c(hodges_list, verbatim_name),
      names_to = "date",
      values_to = "presence"
    ) %>%
    # remove NA rows (absences)
    mutate(presence = ifelse(is.na(presence), 0, 1)) %>%
    filter(presence != 0) %>%
    # remove presence column
    select(-presence) %>%
    # fix dates
    mutate(date = str_replace(date, "Sept", "Sep")) %>%
    mutate(date = as.Date(paste(date, year), format = "%b %d %Y"))
  
  return(df)
}

# define the wrangle_data function for stebs
wrangle_data_stebs <- function(df, year) {
  # convert to df
  df <- df %>% as.data.frame()
  
  # remove blank columns
df <- df %>%
  select(where(~ {
    if (is.numeric(.) || is.logical(.)) {
      !all(is.na(.) | . == 0 | . == FALSE)
    } else {
      TRUE  # Keep non-numeric and non-logical columns
    }
  }))
    
  # apply the function only to the first row
  df[1, ] <- fill_row(as.vector(df[1, ]))
  
  # make a row for the dates
  date_row <- df %>%
    slice(c(1, 2)) %>%
    mutate_all(~ paste(., collapse = " "))
  
  # add it in
  df[1, ] <- date_row[1, ]
  
  # remove daytime collection columns
  df <- df %>%
    select(-matches("D|NT|Trap|day"))
  
  # delete family column
  if (df[3,3] == "ERIOCRANIIDAE" | df[4,3] == "ERIOCRANIIDAE") {
    df <- df %>%
    select(-3)   
  }
  
  # remove columns 4 and 5
  df <- df %>%
    select(-4:-5)
  
  # remove the last 2 columns
  df <- df[, -c((ncol(df) - 1):ncol(df))]
  
  # rename the columns as the first row
  df <- df %>%
    setNames(unlist(df[1, ]))
  
  # define new column names for the species columns
  new_names <- c("hodges", "list", "verbatim_name")
  
  # small edits add up :)
  df <- df %>%
    # rename the species columns
    rename_with(~ new_names, c(1, 2, 3)) %>%
    # remove the first 2 rows using dplyr
    slice(-(1:2)) %>%
    # drop columns where "verbatim_name" is NA
    drop_na(verbatim_name) %>%
    # remove extra text
    filter(!verbatim_name == "Species:",
           !verbatim_name == "species (total)",
           !verbatim_name == "Individuals:",
           !verbatim_name == "Off:",
           !verbatim_name == "On:",
           !verbatim_name == "Temp., deg. F.:",
      !verbatim_name == "Hi Temp",
      !verbatim_name == "Lo Temp"
           ) %>%
    # remove periods from names
    rename_with(~ gsub("\\.", "", .), contains(".")) %>%
    # combine the first two columns into a single column
    mutate(hodges_list = ifelse(is.na(list), hodges, paste0(hodges, ".", list))) %>%
    # remove columns that were just used
    select(-c(hodges, list)) %>%
    # put hodges_list first
    select(hodges_list, everything()) %>%
    # pivot dates longer
    pivot_longer(
      cols = -c(hodges_list, verbatim_name),
      names_to = "date",
      values_to = "presence"
    ) %>%
    # remove NA rows (absences)
    mutate(presence = ifelse(is.na(presence), 0, 1)) %>%
    filter(presence != 0) %>%
    # remove presence column
    select(-presence) %>%
    # fix dates
    mutate(date = str_replace(date, "Sept", "Sep")) %>%
   mutate(date = as.Date(paste(date, year), format = "%b %d %Y"))
  
  return(df)
}

# wrangle data
davis_wrangled <- imap(davis, wrangle_data_davis) %>% 
  list_rbind() %>% drop_na(date)
stebs_wrangled <- imap(stebs, wrangle_data_stebs) %>% 
  list_rbind() %>% drop_na(date)

```

### ALL NAMES

This is John's latest iteration of a name-matching db.

```{r link-across-sites}

# load in all names excel sheet
all_names <- read_excel("../01_raw-data/ALL NAMES.xlsx")

# remove extraneous columns
all_names <- all_names %>%
  select(-c(3, 5, 6, 7, 9, 10), -(11:14), -(16:19))

# vector of new names
col_names <- c(
  "hodges_list",
  "p_no",
  "favored_name",
  "davis_formatted",
  "davis_spreadsheet",
  "stebs_formatted",
  "stebsum_name",
  "is_highlighted"
)

# rename
names(all_names) <- col_names

# remove the first 4 rows; remove if no favored name
all_names <- all_names %>%
  slice(-(1:4)) %>%
  filter(!is.na(favored_name))

# removals
all_names <- all_names %>%
  filter(favored_name != "Aglossa caprealis (Hübner)")


```

### Formatted "favored" names

Double check that every observation has a "favored" name.

```{r favored-names}

# load in list of names
preferred_names_stebs <- read_excel("../01_raw-data/formatted-names-GMH-v2.xlsx")

preferred_names_stebs <- preferred_names_stebs %>%
  # remove the 1st row
  slice(-1) %>%
  # combine the first and second columns
  mutate(hodges_list = ifelse(!is.na(`...2`), 
                              paste0(Hodges, ".", `...2`), 
                              Hodges)) %>%
  # remove columns used to make hodges_list
  select(hodges_list, everything(), -Hodges, -`...2`)

# vector of new names
col_names <- c(
  "hodges_list",
  "p_no",
  "verbatim_name",
  "preferred_name",
  "stebsum_name",
  "old_name1",
  "old_name2",
  "old_name3",
  "old_name4",
  "old_name5" ,
  "old_name6",
  "old_name7",
  "comments"
)

# rename
names(preferred_names_stebs) <- col_names

# remove hodges_list and p_no from preferred_names_stebs
all_names_joined <- preferred_names_stebs %>%
  select(-hodges_list, -p_no) %>%
  rename(old_name9 = stebsum_name) %>%
  full_join(all_names, by = c("preferred_name" = "favored_name"))

# rename 2 columns and then reorder to preferred order
all_names_joined <- all_names_joined %>%
  rename(old_name8 = verbatim_name) %>%
  select(hodges_list, p_no, preferred_name, davis_formatted,
         davis_spreadsheet, stebs_formatted, stebsum_name,
         old_name1, old_name2, old_name3, old_name4, old_name5,
         old_name6, old_name7, old_name8, old_name9)

# get unique names from stebs data
unique_stebs <- stebs_wrangled %>% 
  select(verbatim_name) %>% 
  distinct()

# get unique names from davis data
unique_davis <- davis_wrangled %>% 
  select(verbatim_name) %>% 
  distinct()

# merge unique names together
unique_names <- rbind(unique_davis, unique_stebs) %>% distinct()

# make a database with final columns and a list of old names
all_names_db <- all_names_joined %>%
  select(hodges_list, p_no, preferred_name) %>%
  mutate(old_names = map(1:nrow(.), function(i) {
    c(
      all_names_joined$preferred_name[i],
      all_names_joined$davis_formatted[i],
      all_names_joined$davis_spreadsheet[i],
      all_names_joined$stebs_formatted[i],
      all_names_joined$stebsum_name[i],
      all_names_joined$old_name1[i],
      all_names_joined$old_name2[i],
      all_names_joined$old_name3[i],
      all_names_joined$old_name4[i],
      all_names_joined$old_name5[i],
      all_names_joined$old_name6[i],
      all_names_joined$old_name7[i],
      all_names_joined$old_name8[i],
      all_names_joined$old_name9[i]
    ) %>%
      discard(is.na)
  })) %>% distinct()

# unnest the old_names column so that each old name is in its own row
df_long <- all_names_db %>%
  unnest(old_names)

# join this long dataframe with the unique_names dataframe by old_name
unique_names_with_preferred <- unique_names %>%
  left_join(df_long, by = c("verbatim_name" = "old_names")) %>%
  rename("database_name" = "verbatim_name") %>%
  distinct() %>%
  select(hodges_list, p_no, everything())

```

### Replace database names with preferred names

```{r name-updates}

# update names
stebs_wrangled_updated <- stebs_wrangled %>%
  # remove list number
  select(-hodges_list) %>%
  # left join with preferred names
  left_join(., unique_names_with_preferred, by = join_by(verbatim_name == database_name)) %>%
  # put them in order
  select(hodges_list, p_no, preferred_name, date, verbatim_name) %>%
  # drop observations w/o preferred_name
  drop_na(preferred_name)

# update names
davis_wrangled_updated <- davis_wrangled %>%
  # remove list number
  select(-hodges_list) %>%
  # left join with preferred names
  left_join(., unique_names_with_preferred, by = join_by(verbatim_name == database_name)) %>%
  # put them in order
  select(hodges_list, p_no, preferred_name, date, verbatim_name) %>%
  # drop observations w/o preferred_name
  drop_na(preferred_name)


```

### ⚠️ Add John's confidence levels

```{r john-species-confidence}


```

### Calculate zero years for all species

```{r zero-years}

# make dataframe with preferred name and date only
names_obs_stebs <- stebs_wrangled_updated %>%
  select(preferred_name, date)

names_obs_davis <- davis_wrangled_updated %>%
  select(preferred_name, date)


fill_missing_years <- function(df) {
  # Convert date to year and count observations per species per year
  yearly_counts <- df %>%
    mutate(year = year(as.Date(date))) %>%
    group_by(preferred_name, year) %>%
    summarize(observations = n(), .groups = "drop")
  
  # Get the range of years
  min_year <- min(yearly_counts$year)
  max_year <- max(yearly_counts$year)
  all_years <- min_year:max_year
  
  # Get unique preferred names
  preferred_names <- unique(df$preferred_name)
  
  # Create a dataframe with all combinations of preferred_name and year
  all_combinations <- expand.grid(preferred_name = preferred_names, year = all_years)
  
  # Merge with the yearly counts
  result <- all_combinations %>%
    left_join(yearly_counts, by = c("preferred_name", "year")) %>%
    mutate(day_positives = ifelse(is.na(observations), 0, 1)) %>%
    select(preferred_name, year, day_positives) %>%
    arrange(preferred_name, year) %>%
    rename(matched_name = preferred_name)
  
  return(result)
}

# Apply the function
stebs_zeros <- fill_missing_years(names_obs_stebs) %>% 
  filter(day_positives == 0)

davis_zeros <- fill_missing_years(names_obs_davis) %>% 
  filter(day_positives == 0)

```

### Match names to a backbone

First, I made a list of names, then I exported that list so that I could run a command line script to process the data and tie it back to the GBIF and iNat taxonomy backbones. Then, I created a column based on the sort scores so that I am able to identify which "species" need to be manually checked.

```{r name-match}

# create a function to match names to a backbone
get_names <- function(df, site) {

  # get species names
  taxa <- unique(df$preferred_name)
  
  # construct file name string
  filename <- paste0("../01_raw-data/", site, "-names/", 
                     deparse(substitute(df)), ".txt")
  
  # construct file output string
  fileoutput <- paste0("../01_raw-data/", site, "-names/", 
                       deparse(substitute(df)), "-output.csv")
  
  # write to a text file for use with names verifier
  writeLines(taxa, filename)
  
  # parse names (prefer iNat (180) and GBIF (11))
  system(paste("gnverifier", filename, ">", fileoutput, " -s '11, 180'"))
  
  # read in matched names
  check_names <- read_csv(fileoutput)
  
  # select columns, rename, and join with obs data
  names_df <- check_names %>%
    select(SortScore, ScientificName, MatchedCanonical, 
           TaxonId, DataSourceTitle, ClassificationPath) %>%
    rename(sort_score = SortScore,
           matched_name = ScientificName,
           genus_species = MatchedCanonical,
           taxon_id = TaxonId,
           source = DataSourceTitle) %>%
    # create a column for bad matches or no matches
    mutate(manual_check = ifelse(sort_score < 9 | str_count(genus_species, "\\S+") < 2 | str_detect(matched_name, "\\?"), 1, 0),
           family = str_match(ClassificationPath, "(?:[^|]*\\|){4}([^|]*)")[,2]) %>%
    select(-ClassificationPath) %>%
    right_join(., df, join_by(matched_name ==  preferred_name))
  
  return(names_df)
}

# run function on davis names
davis_full <- get_names(davis_wrangled_updated, "davis")
stebs_full <- get_names(stebs_wrangled_updated, "stebs")

```

### ⚠️ Add checklist number

I joined the Moth Photographer's group checklist numbers to John's data.

```{r pohl-checklist}
#| warning: false

# load in the checklist data
checklist <- read_excel("../01_raw-data/MPG-Taxa_20240311.xlsx") %>%
  rename(
    p_no = `P No`,
    genus_species = Genus_Species,
    common_name = `Common Name`,
    family = Family,
    genus = Genus,
    species = Species,
    synonymy = Synonymy
  ) %>%
select(p_no, MONA, genus_species, common_name, 
       family, genus, species, synonymy)

# update the taxonomy

# if genus_species from stebs_full is in the synonymy column (but not
# necesarily exactly equal), then replace genus_species from the
# data frame stebs_full with genus_species from the df checklist
stebs_full <- stebs_full %>%
  mutate(genus_species = case_when(
    sapply(genus_species, function(gs) {
      # Split the name into its components
      parts <- str_split(gs, " ")[[1]]
      # Create patterns for both binomial and trinomial matches
      patterns <- c(paste0("\\b", paste(parts[1:2], collapse = " "), "\\b"),
                    if (length(parts) > 2)
                      paste0("\\b", paste(parts[1:3], collapse = " "), "\\b")
                    else
                      NULL)
      # Check if any pattern matches
      any(sapply(patterns, function(pat)
        any(str_detect(
          checklist$synonymy, pat
        ))))
    }) ~
      sapply(genus_species, function(gs) {
        parts <- str_split(gs, " ")[[1]]
        patterns <- c(paste0("\\b", paste(parts[1:2], collapse = " "), "\\b"),
                      if (length(parts) > 2)
                        paste0("\\b", paste(parts[1:3], collapse = " "), "\\b")
                      else
                        NULL)
        for (pat in patterns) {
          match_index <- which(str_detect(checklist$synonymy, pat))
          if (length(match_index) > 0)
            return(checklist$genus_species[match_index[1]])
        }
        return(gs)
      }),
    TRUE ~ genus_species
  ))

# join the datasets together
davis_full <- left_join(davis_full, checklist)
stebs_full <- left_join(stebs_full, checklist)

```

### Terraclimate data

I scraped the data from terraclimate for Davis and Cold Canyon locations between 1989 and 2023. I then put the data is a list called `dat` with sublists for Davis and Cold Canyon.

```{r terraclimate}

# make a df with dates
d1 <- as.Date("19580101", "%Y%m%d") # start date
d2 <- as.Date("20231231", "%Y%m%d") # end date

# fill in dates between
dat <- format(seq(d1, d2, by = "month"), "%m %Y") %>% 
  as_tibble() %>%
  separate(value, into = c("month", "year"), sep = " ")

# list for davis and stebs with all dates
dat <- list("davis" = dat, "stebs" = dat)

# for loop for both sites to get data
for (site in c("davis", "stebs")) {
  
  temp_dat = tibble(.rows = 792) # must equal number of months in dataset
  
  if (site == "davis") {
    x = c(-121.78, 38.5569) # Davis long, lat
  } else if (site == "stebs") {
    x = c(-122.0972, 38.51) # Stebs long, lat
  }
  
  # gather all variables of interest
for (var in c("tmax", "tmin", "ppt")) {
  baseurlagg <-
    paste0(
      paste0(
        "http://thredds.northwestknowledge.net:8080/thredds/dodsC/agg_terraclimate_",
        var
      ),
      "_1958_CurrentYear_GLOBE.nc"
    )
  
  nc <- nc_open(baseurlagg)
  lon <- ncvar_get(nc, "lon")
  lat <- ncvar_get(nc, "lat")
  flat = match(abs(lat - x[2]) < 1 / 48, 1)
  latindex = which(flat %in% 1)
  flon = match(abs(lon - x[1]) < 1 / 48, 1)
  lonindex = which(flon %in% 1)
  start <- c(lonindex, latindex, 1)
  count <- c(1, 1, -1)
  
  # read in the full period of record using aggregated files
  data = as.numeric(ncvar_get(nc, varid = var, start = start, count)) %>%
    as_tibble_col(., column_name = var)
  temp_dat = temp_dat %>% add_column(data)
  
}
  temp_dat = temp_dat %>% mutate(dataset = site)
  dat[[site]] = cbind(temp_dat, dat[[site]])
}

TerraClimate <- dat %>% map_dfr(~ as_tibble(.)) %>% filter(year >= 1989 & year < 2024) %>%
  mutate(source = "terraclimate",
         month = as.numeric(month),
         year = as.numeric(year))

```

### Format seasonal weather data

Here, I summarize the weather data by season instead of month:

-   First, I need to make a new column for `water_year`. For example, water year 1998 includes September through December of calendar year 1997

-   Next, I need to group variables by season. Fall (suffix = `_at`) = months 9, 10, 11, winter (`_wt`) = 12, 1, 2, spring (`_sp`) = 3, 4, 5, and summer (`_sm`) = 6, 7, 8.

```{r seasonal-weather}

# function to assign season based on month
get_season <- function(month) {
  case_when(
    month %in% c(9, 10, 11) ~ "at",
    month %in% c(12, 1, 2) ~ "wt",
    month %in% c(3, 4, 5) ~ "sp",
    month %in% c(6, 7, 8) ~ "sm"
  )
}

# create water_year and season columns, remove month column
# set water year column (September to August)
TerraClimate <- TerraClimate %>%
  mutate(
    water_year = ifelse(month %in% c(9, 10, 11, 12), year - 1, year),
    season = get_season(month)
  ) %>%
  select(-month)

# weather data per year
weather <- TerraClimate %>%
  group_by(season, dataset, year) %>%
  summarise(
    tmin = mean(tmin, na.rm = TRUE),
    tmax = mean(tmax, na.rm = TRUE),
    ppt = sum(ppt, na.rm = TRUE)
  )

# make into wider format
weather <- weather %>%
  # pivot the data longer first
  pivot_longer(
    cols = c(tmin, tmax, ppt),
    names_to = "variable",
    values_to = "value"
  ) %>%
  # now create the new column names
  mutate(new_col = paste(variable, season, sep = "_")) %>%
  # pivot wider to get the desired format
  pivot_wider(
    id_cols = c(dataset, year),
    names_from = new_col,
    values_from = value
  )

# join weather data to observation data
stebs_full <- stebs_full %>%
  mutate(month = month(date), 
         year = year(date), 
         season = get_season(month)) %>%
  left_join(weather %>% filter(dataset == "stebs"), by = c("year"))

davis_full <- davis_full %>%
  mutate(month = month(date), 
         year = year(date), 
         season = get_season(month)) %>%
  left_join(weather %>% filter(dataset == "davis"), by = c("year"))
  
```

### Calculate adjusted FDPs

```{r fdps}

# Function to calculate Fractional Day Positives (FDP)
calculate_FDP <- function(data, zeros_data) {
  require(dplyr)
  require(lubridate)

  # Convert date to year and day of year (DOY)
  data <- data %>%
    mutate(year = year(date), DOY = yday(date))

  # Calculate day_positives and collection_events
  day_positives <- data %>%
    group_by(matched_name, year) %>%
    summarise(day_positives = n(), .groups = 'drop') %>%
    rbind(zeros_data)

  collection_events <- data %>%
    group_by(year) %>%
    summarise(collection_events = n_distinct(date), .groups = 'drop')

  # Calculate flight_min and flight_max
  flight_window <- data %>%
    group_by(matched_name) %>%
    summarise(
      flight_min = min(DOY),
      flight_max = max(DOY),
      .groups = 'drop'
    )

  # Calculate adj_collection_events
  adj_collection_events <- data %>%
    left_join(flight_window, by = "matched_name") %>%
    distinct(matched_name, year, date, flight_min, flight_max) %>%
    group_by(matched_name, year) %>%
    summarise(
      adj_collection_events = sum(yday(date) >= flight_min & yday(date) <= flight_max),
      .groups = 'drop'
    )

  # Combine all results
  result <- day_positives %>%
    left_join(collection_events, by = "year") %>%
    mutate(day_negatives = collection_events - day_positives) %>%
    left_join(flight_window, by = "matched_name") %>%
    left_join(adj_collection_events, by = c("matched_name", "year"))

  # Merge results with original data
  data <- data %>%
    select(-DOY) %>%  # Remove the temporary DOY column
    left_join(result, by = c("matched_name", "year"))

  return(data)
}

# List of dataframes to iterate over
data_list <- list(
  stebs = stebs_full,
  davis = davis_full
)

# List of corresponding zeros dataframes
zeros_list <- list(
  stebs = stebs_zeros,
  davis = davis_zeros
)

# Apply the function to each dataframe in the list
results <- purrr::map2(data_list, zeros_list, calculate_FDP)

# Flatten the list by appending data frames and create a new column based on list names
full_data <- purrr::map2_dfr(results, names(results), ~ mutate(.x, dataset = .y))

```

### Add in pest, fire, trail, and SST data

At this juncture, I only want to know if a species is a pest or not. So, I will create a list of all species in the pest data frame. If a binomial is in it, then it will receive a 1 in the column `any_pest` if not it will receive a 0. Trail and fire need only to be joined to the Cold Canyon data.

```{r predictors}

# get fire years 
fire_predictor <- read_csv("../03_processed-data/predictors/fire_predictor.csv")

# get trail years
trail_predictor <- read_csv("../03_processed-data/predictors/trail_predictor.csv")

# get SST predictor
SST_predictor <- read_csv("../03_processed-data/predictors/meiv2.csv")

# get average SST per year and pivot longer
SST_predictor <- SST_predictor %>%
  pivot_longer(everything(), names_to = "year", values_to = "value") %>%
  group_by(year) %>%
  summarize(SST = mean(value)) %>%
  mutate(year = as.numeric(year))

# join fire and trail predictors to stebs data
# filter for dataset == "stebs"
stebs_data <- full_data %>%
  filter(dataset == "stebs")

# perform the joins only for the filtered data
stebs_data <- stebs_data %>%
  left_join(fire_predictor) %>%
  left_join(trail_predictor)  

# combine with the original data, keeping the rows where dataset != "stebs"
full_data <- full_data %>%
  filter(dataset != "stebs") %>%
  bind_rows(stebs_data)

# load pest data
pest_predictor <- read_csv("../03_processed-data/predictors/pest_status.csv")

# reshape the dataframe to long format
long_pest_predictor <- pest_predictor %>%
  pivot_longer(cols = everything(), names_to = "crop", values_to = "presence")

# get unique pest species
unique_pests <- unique(long_pest_predictor[[2]])  # The first column name might vary

# add pest status to full df
full_data  <- full_data %>%
  mutate(any_pest = ifelse(genus_species %in% unique_pests & !is.na(genus_species), 1, 0))

# add SST to full data
full_data <- full_data %>%
  left_join(SST_predictor)

```

### Write final dataset

```{r}

# write data to file
write_csv(full_data, "../03_processed-data/full_data.csv")

```
